### 2015-05-23 docs/report/intro.tex
Corrections up to and including initialization.tex

In the year 2010 it was predicted \cite{itrs} that this tendency would be slowly reaching its end, mostly due to reasons related to heat dissipation. This is the cause why newer computer have a greater number of processors instead of processors with more computing power.

To be able to maximize the performance of these computers, it is paramount that systems professionals can create programs that process data in parallel, through different mechanisms, such as threads or processes executing simultaneously in different processors. In the case of application executed on a single device (e.g. browsers and web pages, applications for mobile devices, desktop apps, server applications) a lot of languages and platforms provide a simple way for programmers to abstract the complexity and coordination required for this type of processing. For example, the .NET platform has the Task Parallel Library (TPL) and Parallel LINQ (PLINQ) \cite{ms-par} libraries that use algorithmic skeletons modelled with high order functions so developers can perform complex operations through a simple API.

On a different note, one of the tendencies that is beggining to take off is the use of EcmaScript \cite{es-web} (also known as Javascript) to create applications that were previously only considered viable in a native environment, such as video games. This has been possible thanks to components like asm.js \cite{asm-web} and WebGL \cite{webgl-web} together with the constant evolution of browsers and JavaScript runtime engines.
In this context, one of the plans of the commitee that develops the language is to provide an API to simplify the processing of data in parallel for version 7 (ES7) of the language, with the goal of keeping up with native applications. Initiatives such as ParallelJS \cite{par-js} or the possibility of taking advantage of SIMD instructions \cite{js-simd} (single instruction multiple data) are some of the options to implement this.

The goal of this professional assignment is to provide an alternative for parallel code eection in EcmaScript, through a library that exposes high order functions to model algorithmic skeletons such as map, filter and reduce. This library will take advantage of Web Workers \cite{w3c-ww} (the mechanism proposed by browsers for parallelism at the time of writing) to allow the simultaneous execution of code in multiple processors.

>>>>

In the year 2010 it was predicted that this tendency would be slowly reaching its end, mostly due to reasons related to heat dissipation. This is the reason why newer computers have a greater number of processors instead of processors with more computing power.

To be able to maximize the performance of these computers, it is paramount that systems professionals can create programs that process data in parallel, through different mechanisms, such as threads or processes executing simultaneously in different processors. In the case of applications executed on a single device (e.g. browsers and web pages, applications for mobile devices, desktop apps, server applications) a lot of languages and platforms provide a simple way for programmers to abstract the complexity and coordination required for this type of processing. For example, the .NET platform has the Task Parallel Library (TPL) and Parallel LINQ (PLINQ) \cite{ms-par} libraries that use algorithmic skeletons modelled with high order functions so developers can perform complex operations through a simple API.

On a different note, one of the tendencies that is beginning to take off is the use of EcmaScript \cite{es-web} (also known as Javascript) to create applications, such as video games, that were previously only considered viable in a native environment. This has been possible thanks to components like asm.js \cite{asm-web} and WebGL \cite{webgl-web} together with the constant evolution of browsers and JavaScript runtime engines.
In this context, one of the plans of the commitee that develops the language is to provide an API to simplify the processing of data in parallel for version 7 (ES7) of the language, with the goal of keeping up with native applications. Initiatives such as ParallelJS \cite{par-js} and River trail \cite{rivertrail} or the possibility of taking advantage of SIMD instructions \cite{js-simd} (single instruction multiple data) are some of the options to implement this.

The goal of this paper is to provide an alternative for parallel code execution in EcmaScript, through a library that exposes high order functions to model algorithmic skeletons such as map, filter and reduce. This library will take advantage of Web Workers \cite{w3c-ww} (the mechanism proposed by browsers for parallelism at the time of writing) to allow the simultaneous execution of code in multiple processors.

### 2015-05-23 docs/report/micro_benchmarks.tex
Corrections up to and including initialization.tex

\section{About Microbenchmmarks}
A common practice to measure the performance of a fragment of code, or to compare it to the performance of some other code fragment is to create a micro benchmark. Throughout this document, there are multiple references to benchmarks of this kind since their results helped us make decisions about how to best implement some specific library features.

@@

The first point is applicable to all programming languages but the second one is only important in engines/VMs that perform such optimizations. Since this document discussed the performance of JavaScript and in particular v8, the second point does affect our benchmarks.

\subsection{A small example\protect\footnote{The example is based on \cite{mraleph-bc}.}}
Let's assume that we want to profile to the following function:
\begin{lstlisting}[caption=Function to benchmark]
function between(a, b, c){
  return a <= b <= c;
}
\end{lstlisting}

@@

The problem with the above is that it does not consider optimizations. Let's assume that the executing engine is smart enough to:
\begin{enumerate}
  \item Inline functions
  \item Detect loop invariants
\end{enumerate}

@@

If it applies \#2 it realizes it does not need to perform the comparison in every iteration so it could extract it outside the loop, resulting in
\begin{lstlisting}[caption=Naive benchmark after evaluating loop invariants]
// begin setup
var a = Math.round(Math.random()*100);
var b = Math.round(Math.random()*100) + a;
var c = Math.round(Math.random()*100) + b;
//end setup
//start benchmark
console.time('between');
var j = a <= b && b <= c;
for (var i = 0; i < 100000; i++){
}
console.timeEnd('between');
//end benchmark
\end{lstlisting}

Alternatively it could notice that the variable \code{j} is not being used so it could just throw the line away. In any case, we would just be measuring how long it takes for JavaScript to execute the loop, which is not what we originally intended to do. To verify that is not the case the tool IRHydra\cite{ir-hydra} can be used , which \textit{can display intermediate representations used by V8}. As shown in the following figure, v8 inline the function (the blue chevron marks that) but the loop invariant was not detected and optimized:
\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{naive_benchmark_ir_hydra}
  \caption{Optimized benchmark code with inlined function.}
\end{figure}

\note{All benchmarks are performed using two different versions of Google Chrome (40.0.2214 and 41.0.2272) to be able to get an idea of the evolution of each feature.}

>>>>

\section{About Microbenchmmarks}
A common practice to measure the performance of a fragment of code, or to compare it to the performance of some other code fragment is to create a micro benchmark. Throughout this document, there are multiple references to benchmarks of this kind, since their results helped us make decisions about how to best implement some specific library features.

@@

The first point is applicable to all programming languages but the second one is only important in engines/VMs that perform such optimizations. This document discusses the performance of JavaScript and in particular v8, so the second point does affect our benchmarks.

\subsection{A small example\protect\footnote{The example is based on \cite{mraleph-bc}.}}
Let's assume that we want to profile the following function:
\begin{lstlisting}[caption=Function to benchmark]
function between(a, b, c){
  return a <= b <= c;
}
\end{lstlisting}

@@

The problem with the above benchmark is that it does not consider optimizations. Let's assume that the executing engine is smart enough to:
\begin{enumerate}
  \item Inline functions
  \item Detect loop invariants
\end{enumerate}

@@

When it applies \#2, it realizes it does not need to perform the comparison in every iteration so it could extract it outside the loop, resulting in
\begin{lstlisting}[caption=Naive benchmark after evaluating loop invariants]
// begin setup
var a = Math.round(Math.random()*100);
var b = Math.round(Math.random()*100) + a;
var c = Math.round(Math.random()*100) + b;
//end setup
//start benchmark
console.time('between');
var j = a <= b && b <= c;
for (var i = 0; i < 100000; i++){
}
console.timeEnd('between');
//end benchmark
\end{lstlisting}

Alternatively it could notice that the variable \code{j} is not being used so it could just throw the line away. In any case, we would just be measuring how long it takes for JavaScript to execute the loop, which is not what we originally intended to do. To verify that is not the case the tool IRHydra\cite{ir-hydra} can be used , which \textit{can display intermediate representations used by V8}. As shown in the following figure, v8 inlined the function (the blue chevron marks that) but the loop invariant was not detected and optimized:
\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{naive_benchmark_ir_hydra}
  \caption{Optimized benchmark code with inlined function.}
\end{figure}

\subsection{Details}
\label{sub-sec:serialization-and-transference}
The benchmarks are executed using two different machines:

\textbf{Machine 1}
\begin{itemize}
  \item Mac Book Pro Retina
  \item Intel Core i7 2.8GHz
  \item 16 GB RAM DDR3 1600 MHz
  \item SSD 250 GB
  \item Chrome 41.0.2272
\end{itemize}

\textbf{Machine 2}
\begin{itemize}
  \item Mac Book Air
  \item Intel Core i5 1.8GHz
  \item 4 GB RAM DDR3 1600 MHz
  \item SSD 128 GB
  \item Chrome 43.0.2357
\end{itemize}
